### 2025-11-19 paper/reviews/iclr2026/response.md

# Response to reviewers

We thank our three reviewers for their useful feedback and suggestions for improving the paper. We address
specific comments of each reviewer below, and then summarise our planned revisions to the paper.

## Planned revisions to paper

### 2025-11-19 paper/reviews/iclr2026/response.md

# Response to reviewers

We thank our three reviewers for their useful feedback and suggestions for improving the paper. We address
specific comments of each reviewer below, and then summarise our planned revisions to the paper.

## Reviewer AyvP

_Restriction to single papers._ Nothing about our design restricts it to single papers, although for our
evaluation we have only considered excerpts from single papers. A multi-paper setup is an interesting use case
and one we will consider for future work.

_Checks for more complex claims cannot be generated with LLMs. Even for the simple cases, a human has to
verify the generated links._ Keeping a human in the loop for verification purposes seems important from a
trust point of view. However, there do seem to be some likely ways this could be improved: in particular for
cases where an LLM is unable to generate a full solution in one step, we envisage a more interactive workflow
similar to working with Copilot.

The most important consideration here

### 2025-11-19 paper/reviews/iclr2026/response.md

We thank our three reviewers for their useful feedback and suggestions for improving the paper. We address
specific comments of each reviewer below, and then summarise our planned revisions to the paper.

## Reviewer AyvP

_Restriction to single papers._ Nothing about our design restricts it to single papers, although for our
evaluation we have only considered excerpts from single papers. A multi-paper setup is an interesting use case
and one we will consider for future work.

_Checks for more complex claims cannot be generated with LLMs. Even for the simple cases, a human has to
verify the generated links._ Keeping a human in the loop for verification purposes seems important from a
trust point of view. However, there do seem to be some ways this could be improved: in particular for cases
where an LLM is unable to generate a full solution in one step, we envisage a more interactive workflow
similar to working with Copilot. For the manual validation, automated testing should be able to generate
counterexamples. But the key point is that for a relatively small amount of author effort, every _reader_
benefits. So some kind of hybrid system where the "easy" cases are fully or semi-automated and the more
complex cases are constructed by hand (with a Copilot-like assistant) seems the most plausible.

### 2025-11-19 paper/reviews/iclr2026/response.md

We thank our three reviewers for their useful feedback and suggestions for improving the paper. We address
specific comments of each reviewer below, and then summarise our planned revisions to the paper.

## Reviewer AyvP

_Restriction to single papers._ Nothing about our design restricts it to single papers, although for our
evaluation we have only considered excerpts from single papers. A multi-paper setup is an interesting use case
and one we will consider for future work.

_Checks for more complex claims cannot be generated with LLMs. Even for the simple cases, a human has to
verify the generated links._ Keeping a human in the loop for verification purposes seems important from a
trust point of view. For the simple cases, where it may seem unfortunate that author verification is still
needed, the key point here is that a small amount of author effort translates into a significant _reader_
benefit (since there are many more readers than authors). In future work we anticipate that automated testing
can reduce the author burden in this cases by automatically generating counterexamples.

More the more complex cases where current LLMs are unable to generate a full solution in one step, we envisage
a more interactive workflow similar to working with Copilot. some kind of hybrid system where the "easy" cases
are fully or semi-automated and the more complex cases are constructed by hand (with a Copilot-like assistant)
seems the most plausible.

### 2025-11-19 paper/reviews/iclr2026/response.md

# Response to reviewers

We thank our three reviewers for their useful feedback and suggestions for improving the paper. We address
specific comments of each reviewer below, and then summarise our planned revisions to the paper.

## Reviewer AyvP

_Restriction to single papers._ Nothing about our design restricts it to single papers, although for our
evaluation we have only considered excerpts from single papers. A multi-paper setup is an interesting use case
and we will consider for a real-world tool.

_Checks for more complex claims cannot be generated with LLMs. Even for the simple cases, a human has to
verify the generated links._ Keeping a human in the loop for verification purposes seems important from a
trust point of view. For the simple cases, where it may seem unfortunate that author verification is still
needed, a small amount of author effort nevertheless translates into a significant _reader_ benefit (since
there are many more readers than authors). In future work we anticipate automated testing as a way of reducing
the author burden by automatically exercising the generated code.

For the more complex cases, where current LLMs are unable to generate a full solution in one step, we envisage
a more interactive workflow, similar to working with Copilot. For the present paper we are focused on
exploring the limits of current LLM capabilities and validating the basic idea.

### 2025-11-20 paper/reviews/iclr2026/response.md

## Reviewer AyvP

_Restriction to single papers._ Nothing about our design restricts it to single papers, although for our
evaluation we have only considered excerpts from single papers. A multi-paper setup is an interesting use case
and we will consider for a real-world tool.

_Checks for more complex claims cannot be generated with LLMs. Even for the simple cases, a human has to
verify the generated links._ Keeping a human in the loop for verification purposes seems important from a
trust point of view. For the simple cases, where it may seem unfortunate that author verification is still
needed, a small amount of author effort nevertheless translates into a significant _reader_ benefit (since
there are many more readers than authors). In future work we anticipate automated testing as a way of reducing
the author burden by automatically exercising the generated code.

For the more complex cases, where current LLMs are unable to generate a full solution in one step, we envisage
a more interactive workflow, similar to working with Copilot. For the present paper we are focused on
exploring the limits of current LLM capabilities and validating a simpler architecture with less interaction.

## Reviewer Bezq

_Re. subset of SciGen, could you provide details on the selection logic, sampling criteria, and exact number
of examples used in evaluation?_

For the submitted version we relied on a small subset of SciGen, with only XX example replacement tasks.

### 2025-11-20 paper/reviews/iclr2026/response.md

_Checks for more complex claims cannot be generated with LLMs. Even for the simple cases, a human has to
verify the generated links._ Keeping a human in the loop for verification purposes seems important from a
trust point of view. For the simple cases, where it may seem unfortunate that author verification is still
needed, a small amount of author effort nevertheless translates into a significant _reader_ benefit (since
there are many more readers than authors). In future work we anticipate automated testing as a way of reducing
the author burden by automatically exercising the generated code.

For the more complex cases, where current LLMs are unable to generate a full solution in one step, we envisage
a more interactive workflow, similar to working with Copilot. For the present paper we are focused on
exploring the limits of current LLM capabilities and validating a simpler architecture with less interaction.

## Reviewer Bezq

_Can authors report statistics about the manual validation process?_

Unfortunately a user study is out-of-scope for this work; see [...]

_Re. the subset of SciGen used, could you provide details on the selection logic, sampling criteria, and exact
number of examples used in evaluation?_

[Number of example papers + replacement problems within each of those]

The paper states that experiments were conducted on “a subsample of the SciGen dataset” (line 367), but it remains unclear how this subset was chosen. Could the authors provide more details on the selection logic, sampling criteria, and the exact number of examples used in evaluation?

It seems this framework rely on predefined helper routines such as trendWord or growShrink, which encode semantic logic that the model merely invokes rather than learns. Moreover, the large performance gap between target-value-sharing (74.9%) and no-target (57.1%) suggests potential reliance on implicit answer leakage rather than actual LLM reasoning. Could the authors conduct an ablation study removing or varying these helper components to determine how much of the system’s success derives from the LLM’s own reasoning versus predefined components?

### 2025-11-20 paper/reviews/iclr2026/response.md

## Reviewer Bezq

_Can authors report statistics about the manual validation process?_

Unfortunately a user study is out-of-scope for this work; see [...]

_Re. the subset of SciGen used, could you provide details on the selection logic/sampling criteria and exact
number of examples used in the evaluation?_

@@

[Number of example papers + replacement problems within each of those]

However we have increased the size of the dataset in two ways:
- for measuring correctness only, we have added automation scripts to allow use to process the whole of SciGen (400+ examples)
- for measuring

### 2025-11-21 paper/reviews/iclr2026/response.md

For the more complex cases, where current LLMs are unable to generate a full solution in one step, we envisage
a more interactive workflow, similar to working with Copilot. For the present paper we are focused on
exploring the limits of current LLM capabilities and validating a simpler architecture with less interaction.

## Reviewer Bezq

_Can authors report statistics about the manual validation process?_

Unfortunately a user study is out of scope for this work. We plan to do that next now that we have a good
handle on the required LLM competencies.

### 2025-11-21 paper/reviews/iclr2026/response.md

# Response to reviewers

We thank our three reviewers for their useful feedback and suggestions for revising the paper. We address specific comments of each reviewer below, and then summarise how we plan to improve the paper.

@@

## Reviewer AyvP

_Restriction to single papers._ Nothing about our design restricts it to single papers, although for our evaluation we have only considered excerpts from single papers. A multi-paper setup is an interesting use case and one we will consider for a real-world tool.

_Checks for more complex claims cannot be generated with LLMs. Even for the simple cases, a human has to verify the generated links._ Keeping a human in the loop for verification purposes seems important from a trust point of view. For the simple cases, where it may seem unfortunate that author verification is still needed, a small amount of author effort nevertheless translates into a significant _reader_ benefit (since there are many more readers than authors). In future work we anticipate providing automated testing as a way of reducing author burden by automatically exercising the generated code.

For the more complex cases, where current LLMs are unable to generate a full solution in one step, our current design should be able to form the basis of a more interactive workflow, similar to working with Copilot. For the present paper we are focused on exploring the limits of current LLM capabilities and validating a simpler architecture with less interaction.

### 2025-11-21 paper/reviews/iclr2026/response.md

For the more complex cases, where current LLMs are unable to generate a full solution in one step, our current design should be able to form the basis of a more interactive workflow, similar to working with Copilot. For the present paper we are focused on exploring the limits of current LLM capabilities and validating a simpler architecture with less interaction.

## Reviewer Bezq

_Can authors report statistics about the manual validation process?_

A proper user study with the necessary experimental setup is out of scope for this work, but certainly something we plan to do as the next step, alongside implementing a more complete interactive workflow. For this paper, our main focus is on a basic agent-based architecture for interpretation synthesis and a solid evaluation of of the required LLM competencies.

### 2025-11-21 paper/reviews/iclr2026/response.md

It seems this framework rely on predefined helper routines such as trendWord or growShrink, which encode semantic logic that the model merely invokes rather than learns. Moreover, the large performance gap between target-value-sharing (74.9%) and no-target (57.1%) suggests potential reliance on implicit answer leakage rather than actual LLM reasoning. Could the authors conduct an ablation study removing or varying these helper components to determine how much of the system’s success derives from the LLM’s own reasoning versus predefined components?

For the submitted version we relied on a small subset of SciGen, with only XX example replacement tasks.

## Planned revisions to paper

- Clarify that manually annotated subsets of SciGen are randomly chosen, and carry out additional manual annotation to grow subset from 2% to 5%
- Evaluate SuggestionAgent performance against manual annotations
- Extend our correctness evaluation (Fig. 6(b)) to full SciGen dataset, using SuggestionAgent annotations

### 2025-11-21 paper/reviews/iclr2026/response.md

_Can authors report statistics about the manual validation process?_

A proper user study with the necessary experimental setup is out of scope for this work, but certainly something we plan to do as the next step, alongside implementing a more complete interactive workflow. For this paper, our main focus is on a basic agent-based architecture for interpretation synthesis and a solid evaluation of of the required LLM competencies.

_Re. the subset of SciGen used, could you provide details on the selection logic/sampling criteria and exact number of examples used in the evaluation?_

For the submitted version we were only able to manually annotated a 2% subset of the 400 SciGen issues, with only XX example replacement tasks.

@@

The paper states that experiments were conducted on “a subsample of the SciGen dataset” (line 367), but it remains unclear how this subset was chosen. Could the authors provide more details on the selection logic, sampling criteria, and the exact number of examples used in evaluation?

It seems this framework rely on predefined helper routines such as trendWord or growShrink, which encode semantic logic that the model merely invokes rather than learns. Moreover, the large performance gap between target-value-sharing (74.9%) and no-target (57.1%) suggests potential reliance on implicit answer leakage rather than actual LLM reasoning. Could the authors conduct an ablation study removing or varying these helper components to determine how much of the system’s success derives from the LLM’s own reasoning versus predefined components?

## Planned revisions to paper

- Clarify that manually annotated subsets of SciGen are randomly chosen; add additional manual annotations to grow subset from 2% to 5%
- Evaluate SuggestionAgent performance against manual annotations
- Extend our correctness evaluation (Fig. 6(b)) to full SciGen dataset, using SuggestionAgent annotations

We will include our manually annotated subset as supplementary material and release it as an open source benchmark.

### 2025-11-21 paper/reviews/iclr2026/response.md

_Can authors report statistics about the manual validation process?_

A proper user study with the necessary experimental setup is out of scope for this work, but certainly something we plan to do as the next step, alongside implementing a more complete interactive workflow. For this paper, our main focus is on a basic agent-based architecture for interpretation synthesis and a solid evaluation of of the required LLM competencies.

_Re. the subset of SciGen used, could you provide details on the selection logic/sampling criteria and exact number of examples used in the evaluation?_

For the submitted version we were only able to manually annotate 9 randomly-chosen SciGen issues (approximately 2% of the total), resulting in XX labelled replacement tasks. We have drawn a second (larger) random subsample (5% of the total) for manual annotation and will report on these in the final version of the paper. We have also added an automation pipeline to process every SciGen example using the Suggestion Agent. By evaluating the performance of the SuggestionAgent against the manual annotations to estimate the noise rate, we are now able to evaluate the InterpretationAgent using the entirety of the SciGen data set, with confidence intervals adjusted to incorporate estimated noise.

_The framework rely on predefined helper routines such as trendWord or growShrink_. Could the authors conduct an ablation study removing or varying these helper components?

This is a good suggestion; we will do this and also consider other ablations studies (in addition to excluding the target string). It is also worth emphasising that a fixed/standardised set of definitions (for a given paper or research community, say) might also be helpful, but your point remains valid.

_Large performance gap between target-value-sharing (74.9%) and no-target (57.1%) suggests potential reliance on implicit answer leakage rather than actual LLM reasoning._

@@

## Planned revisions to paper

- Clarify manually-annotated subsets of SciGen are randomly chosen; add additional manual annotations to grow subset from 2% to 5%
- Evaluate SuggestionAgent performance against manual annotations
- Extend our correctness evaluation (Fig. 6(b)) to full SciGen dataset, using SuggestionAgent annotations (and adjusted for noise)

We will include our manually annotated subset as supplementary material and release it as an open source benchmark.

### 2025-11-21 paper/reviews/iclr2026/response.md

For the more complex cases, where current LLMs are unable to generate a full solution in one step, our current design should be able to form the basis of a more interactive workflow, similar to working with Copilot. For the present paper we are focused on exploring the limits of current LLM capabilities and validating a simpler architecture with less interaction.

## Reviewer Bezq

_Can authors report statistics about the manual validation process?_

A proper user study with the necessary experimental setup is out of scope for this work, but certainly something we plan to do as the next step, alongside implementing a more complete interactive workflow. For this paper, our main focus is on a basic agent-based architecture for interpretation synthesis and a solid evaluation of the required LLM competencies.

@@

_Re. the subset of SciGen used, could you provide details on the selection logic/sampling criteria and exact number of examples used in the evaluation?_

For the submitted version we were only able to manually annotate 9 randomly-chosen SciGen issues (approximately 2% of the total), resulting in XX labelled replacement tasks. We have drawn a second (larger) random subsample (5% of the total) for manual annotation and will report on this in the final version of the paper. We have also added an automation pipeline to process every SciGen example using the Suggestion Agent, and will evaluate SuggestionAgent performance using manual annotations to estimate the noise rate. We are now able to evaluate the InterpretationAgent using the entirety of the SciGen data set, with confidence intervals adjusted to incorporate estimated noise.

@@

_The framework rely on predefined helper routines such as trendWord or growShrink_. Could the authors conduct an ablation study removing or varying these helper components?

This is a good suggestion; we will do this and also consider other ablations studies (in addition to excluding the target string). It is worth emphasising that a fixed/standardised set of terminological definitions (for a given paper or research community, say) might be helpful (or even important in some domains, such as IPCC Summary for Policymaker reports), but your point remains valid.

@@

_Large performance gap between target-value-sharing (74.9%) and no-target (57.1%) suggests potential reliance on implicit answer leakage rather than actual LLM reasoning._

## Reviewer VRkd

> Many of the niche details, like Table 1, are perhaps unnecessary for the main paper

We would like to keep Table 1 in the body of the paper to be explicit about the specific linguistic/quantitative LLM competencies we evaluate in the paper. However it may be possible to streamline the figure a bit. We will also make a pass over the paper for other details that might be spurious for our audience here.

@@

## Planned revisions to paper

- Clarify manually-annotated subsets of SciGen are randomly chosen; add additional manual annotations to grow subset from 2% to 5%. We will include this data set as supplementary material and release it as an open source benchmark.
- Evaluate SuggestionAgent performance against manual annotations; estimate noise in SuggestionAgent labelling.
- Extend our correctness evaluation (Fig. 6(b)) to full SciGen dataset, using SuggestionAgent annotations (and adjusted for noise)

### 2025-11-21 paper/reviews/iclr2026/response.md

For the more complex cases, where current LLMs are unable to generate a full solution in one step, our current design should be able to form the basis of a more interactive workflow, similar to working with Copilot. For the present paper we are focused on exploring the limits of current LLM capabilities and validating a simpler architecture with less interaction.

## Reviewer Bezq

_Can authors report statistics about the manual validation process?_

A proper user study with the necessary experimental setup is out of scope for this work, but certainly something we plan to do as the next step, alongside implementing a more complete interactive workflow; we recognise there is a risk (as with all AI-generated code) of simply shifting author workload from writing to debugging. For this paper, our main focus is on a basic agent-based architecture for interpretation synthesis and a solid evaluation of the required LLM competencies.

@@

_The framework rely on predefined helper routines such as trendWord or growShrink_. Could the authors conduct an ablation study removing or varying these helper components?

This is a good suggestion; we will do this and also consider other ablations studies (in addition to excluding the target string). It is worth emphasising that a fixed/standardised set of terminological definitions (for a given paper or research community, say) might be helpful (or even important in some domains, such as IPCC Summary for Policymaker reports), but your point remains valid.

_Large performance gap between target-value-sharing (74.9%) and no-target (57.1%) suggests potential reliance on implicit answer leakage rather than actual LLM reasoning._

In response to additional points:

_No analysis of statistical variance beyond standard deviation._ We are now reporting using box plots to give
a better sense of the underlying distribution, annotated with the number of samples.

### 2025-11-21 paper/reviews/iclr2026/response.md

_The framework rely on predefined helper routines such as trendWord or growShrink_. Could the authors conduct an ablation study removing or varying these helper components?

This is a good suggestion; we will do this and also consider other ablations studies (in addition to excluding the target string). It is worth emphasising that a fixed/standardised set of terminological definitions (for a given paper or research community, say) might be helpful (or even important in some domains, such as IPCC Summary for Policymaker reports), but your point remains valid.

_Large performance gap between target-value-sharing (74.9%) and no-target (57.1%) suggests potential reliance on implicit answer leakage rather than actual LLM reasoning._

The target string is present in one of our main use cases (where the reader or author is trying to retrofit an interpretation to text already written), so this doesn't qualify as answer leakage per se. Counterfactual testing, which is possible for manually annotated solutions, exposes whether some kind of computational reasoning has actually happened. We will improve our reporting on this (RQ2).

To respond to your additional points:

_No analysis of statistical variance beyond standard deviation._ We have changed our figures to use box plots for a better sense of the underlying distribution, annotated with the number of samples.

### 2025-11-21 paper/reviews/iclr2026/response.md

_No analysis of statistical variance beyond standard deviation._ We have changed our figures to use box plots for a better sense of the underlying distribution, annotated with the number of samples.

## Reviewer VRkd

> Many of the niche details, like Table 1, are perhaps unnecessary for the main paper

We prefer to keep Table 1 in the body of the paper, to be explicit about the specific linguistic/quantitative LLM competencies evaluated in the paper; however it should be possible to streamline the figure a bit. We will also make a pass over the paper for other details that might be spurious for the ICLR audience.

@@

## Planned revisions to paper

- Clarify manually-annotated subsets of SciGen are randomly chosen; add additional manual annotations to grow subset from 2% to 5%. We will include this data set as supplementary material and release it as an open source benchmark.
- Evaluate SuggestionAgent performance against manual annotations; estimate noise in SuggestionAgent labelling.
- Extend our correctness evaluation (Fig. 6(b)) to full SciGen dataset, using SuggestionAgent annotations (and adjusted for noise)
- Expand the treatment of RQ2

### 2025-11-21 paper/reviews/iclr2026/response.md

_Large performance gap between target-value-sharing (74.9%) and no-target (57.1%) suggests potential reliance on implicit answer leakage rather than actual LLM reasoning._

The target string is present in one of our main use cases (where the reader or author is trying to retrofit an interpretation to text already written), so this doesn't qualify as answer leakage per se. Counterfactual testing, which is possible for manually annotated solutions, exposes whether some kind of computational reasoning has actually happened. We will improve our reporting on this (RQ2).

To respond to your additional points:

_No analysis of statistical variance beyond standard deviation._

Our figures now use box plots for a better sense of the underlying distribution, annotated with the number of samples.

_Assess generalization to unseen writing styles or datasets._

In our experiments the LLM was not fine-tuned or prompted with SciGen examples other than as part of the test cases themselves, so all the SciGen examples were acting as out-of-distribution problems.

## Reviewer azjF

_More engineering-oriented, with relatively limited academic research value._

_Without comparisons to other baseline methods, it is difficult to determine whether this system represents the optimal solution (though, as a pioneering work, more in-depth ablation studies could be considered)._

Doing something new does indeed make comparison to baseline methods difficult. However we agree more in-depth ablation studies would be appropriate. We are investigating removing the predefined helper routines (see response to reviewer Bezq above); we will consider other ablation possibilities and include those in our reporting on RQ1 and RQ2.

@@

## Reviewer VRkd

> Isn't this supposed to be an HCI paper? To me, there isn't any technical contribution for an ML venue beyond getting the system implemented.
> Since "human" is a really central part in this system, it would be nice to get evaluated by actual human/users too.

The natural next step of this work will include a Copilot-like full-featured tool, which will enable a full user study. This will likelygo to venue like CHI or IUI, as suggested. For this paper, we aim to validate key components of such a tool, including:

- competency evaluation for LLMs (which will substantially improve for this paper)
- proof-of-concept agentic architecture, including support for compiler-in-the-loop and manual validation (both important for this task)

We believe that this is of interest to the ML community, at least from NLP and tool architecture points of view, although we concede that the evaluation in the submitted version of the paper was in need further work, and we plan to substantially address that in the final revision.

### The evaluation, from an LLM/NLP standpoint, is perhaps not enough. It would be nice to show if the system could do well on some documented NLP tasks (like scientific literature?) to begin with.

### 2025-11-21 paper/reviews/iclr2026/response.md

## Reviewer azjF

_More engineering-oriented, with relatively limited academic research value._

_Without comparisons to other baseline methods, it is difficult to determine whether this system represents the optimal solution (though, as a pioneering work, more in-depth ablation studies could be considered)._

The lack of directly related prior work does indeed make comparison to baseline methods difficult. We agree more in-depth ablation studies would be appropriate; we will investigate removing the predefined helper routines (see response to reviewer Bezq above), along with other other similar experiments, and include those in our reporting on RQ1 and RQ2.

@@

## Reviewer VRkd

> Isn't this supposed to be an HCI paper? To me, there isn't any technical contribution for an ML venue beyond getting the system implemented.
> Since "human" is a really central part in this system, it would be nice to get evaluated by actual human/users too.

The natural next step of this work will include a Copilot-like full-featured tool, which will enable a full user study. This will likely go to venue like CHI or IUI, as suggested. For this paper, our aim to validate key components of such a tool, including:

- competency evaluation for LLMs (which we will substantially improve for this paper)
- proof-of-concept agentic architecture, including support for compiler-in-the-loop and manual validation (both important for this task)

### 2025-11-21 paper/reviews/iclr2026/response.md

_Assess generalization to unseen writing styles or datasets._

In our experiments the LLM was not fine-tuned or prompted with SciGen examples other than as part of the test cases themselves, so all the SciGen examples were acting as out-of-distribution problems.

## Reviewer azjF

_More engineering-oriented, with relatively limited academic research value._ It is an engineering-oriented paper, however we address important research questions around how such an agent-based tool might be designed, how currently LLM competencies meet the needs of such a design, and how manual and automatic verification can be integrated and assessed.

_Without comparisons to other baseline methods, it is difficult to determine whether this system represents the optimal solution (though, as a pioneering work, more in-depth ablation studies could be considered)._ The lack of directly related prior work does indeed make comparison to baseline methods difficult. We agree more in-depth ablation studies would be appropriate; we will investigate removing the predefined helper routines (see response to reviewer Bezq above), along with other other similar experiments, and include those in our reporting on RQ1 and RQ2.

@@

## Reviewer VRkd

_Isn't this supposed to be an HCI paper? To me, there isn't any technical contribution for an ML venue beyond getting the system implemented. Since "human" is a really central part in this system, it would be nice to get evaluated by actual human/users too._ The natural next step of this work will include a Copilot-like full-featured tool, which will enable a full user study. And indeed this would likely go to venue like CHI or IUI. For this paper, our aim to validate key components of such a tool, including:

@@

- competency evaluation for LLMs (which we will substantially improve for this paper)
- proof-of-concept agentic architecture, including support for compiler-in-the-loop and manual validation (both important for this task)

We believe that this is of interest to the ML community, at least from NLP and tool architecture points of view, although we concede that the evaluation in the submitted version of the paper was in need further work. We discuss elsewhere how we plan to address that in the final revision.

@@

### The evaluation, from an LLM/NLP standpoint, is perhaps not enough. It would be nice to show if the system could do well on some documented NLP tasks (like scientific literature?) to begin with.

_Many of the niche details, like Table 1, are perhaps unnecessary for the main paper._ We prefer to keep Table 1 in the body of the paper, to be explicit about the specific linguistic/quantitative LLM competencies evaluated in the paper; however it should be possible to streamline the figure a bit. We will also make a pass over the paper for other details that might be spurious for the ICLR audience.

### 2025-11-21 paper/reviews/iclr2026/response.md

# General Response to Reviewers
[Post as initial comment, then one comment per reviewer, as below]

We thank all the reviewers for their time and expertise in evaluating our submission. We are grateful for the positive feedback, including the recognition of [..] contribution (Reviewiers ???), the acknowledgment that [..] well-motivated (Reviewers ??), and the appreciation of (Reviewers ??). The constructive feedback provided by the reviewers has been invaluable in improving the paper. We have provided detailed responses to each reviewer's comments individually and outline below the key results from new experiments conducted in response to multiple reviewer requests.

## Planned Revisions and New Experiments

# Response to Reviewer AyvP

@@

_Isn't this supposed to be an HCI paper? To me, there isn't any technical contribution for an ML venue beyond getting the system implemented. Since "human" is a really central part in this system, it would be nice to get evaluated by actual human/users too._ The natural next step of this work will include a Copilot-like full-featured tool, which will enable a full user study. And indeed this would likely go to venue like CHI or IUI. For this paper, our aim to validate key components of such a tool, including:

- competency evaluation for LLMs (which we will substantially improve for this paper)
- proof-of-concept agentic architecture, including support for compiler-in-the-loop and manual validation (both important for this task)

We believe that this is of interest to the ML community, at least from NLP and tool architecture points of view, although we concede that the evaluation in the submitted version of the paper was in need further work. We discuss elsewhere how we plan to address that in the final revision.

_The evaluation, from an LLM/NLP standpoint, is perhaps not enough. It would be nice to show if the system could do well on some documented NLP tasks (like scientific literature?) to begin with._

### 2025-11-21 paper/reviews/iclr2026/response.md

[Post each of the top-level sections below a separate comment, with title as given]

# General Response to Reviewers

We thank all reviewers for their time and expertise in evaluating our submission. We are grateful for the positive feedback, including the recognition that the proposed system addresses an important research direction (Reviewers AyvP, Bezq and VRkd), the appreciation that the exploration of a connection between LLM-based code synthesis and data provenance systems is novel (Reviewer Bezq) and that our experimental results, while in need of some further work, show that the approach looks feasible in practice (Reviewers ??). The constructive feedback provided by the reviewers has been invaluable in understanding how to improve the evaluation. We have provided detailed responses to each reviewer individually and outline below how we plan to further revise the paper and the new experiments conducted in response to multiple reviewer requests.

### 2025-11-21 paper/reviews/iclr2026/response.md

## Planned Revisions and New Experiments

- Clarify manually-annotated subsets of SciGen are randomly chosen; add additional manual annotations to grow subset from 2% to 5%. We will include this data set as supplementary material and release it as an open source benchmark.
- Evaluate SuggestionAgent performance against manual annotations; estimate noise in SuggestionAgent labelling.
- Extend our correctness evaluation (Fig. 6(b)) to full SciGen dataset, using SuggestionAgent annotations (and adjusted for noise)
- Expand the treatment of RQ2

# Response to Reviewer AyvP

> - The proposed system still seems to be in early stages with respect to what can be verified, which limits its usefulness in practice. First, as far as I understand, it is limited to single papers and does not allow to check claims a paper makes about results presented in another paper. This is where a system like the proposed would be most useful

Thank you for the suggestion; a multi-paper setup is indeed an interesting use case. Although for our evaluation we only considered excerpts from single papers, nothing about our design restricts it to such, so we will consider this for a real-world tool.

> - Second, the claims that can be checked seem to be quite simple (checks for more complex claims cannot be generated with LLMs) and thus easy to check manually. Even then, a human has to verify all generated links, which in at least some cases are wrong. This begs the question of whether a manual system where only larger and more important claims are annotated would not be more useful in practice.

Keeping a human in the loop for verification purposes seems important from a trust point of view. For the simple cases, where it may seem unfortunate that author verification is still needed, a small amount of author effort nevertheless translates into a significant _reader_ benefit (since there are many more readers than authors). In future work we anticipate providing automated testing as a way of reducing author burden by automatically exercising the generated code.

### 2025-11-21 paper/sec/authoring-workflow.tex
Fix link to section 4.
Break down by RQ1 and RQ2.

\item \textbf{High-level Authoring workflow.} The system then enters the human-in-the loop authoring workflow
shown in Figure~\ref{fig:interpretation-agent}, where the author interacts with the \InterpretationAgent. The
system waits for the author to select a fragment of text $s$ to interpret (perhaps previously highlighted by
the \SuggestionAgent). The system then attempts to generate a candidate Fluid expression $e$ using the
closed-loop synthesis step (3) below. If code synthesis succeeds with an expression $e$, the system proceeds
to the manual validation step (4) below. If the synthesis step fails with no expression, no remedial action is
possible; this is considered an unsuccessful path through the workflow and returns the system to the entry
state. Otherwise the synthesis step produces an expression $e$ which evaluates to a mismatched string $s' \neq
s$ outcome, and the user can choose to manually abort and return to the entry state, or optionally to
\emph{revise the goal}, replacing $s$ with $s'$ in the target document and retaining $e$ as the candidate
expression. This is intended to cover the situation where the author has made a claim which is
\emph{incorrect}, and the data set and surrounding natural language have led the LLM to synthesise an
expression which generates a different value from the one specified by the user.

@@

\item \textbf{Manual validation step.} Once a candidate expression has been generated, the system replaces the
selected substring $s$ with the interpolation expression \kw{\{$e$\}}, creating a new (but only tentative)
document configuration. The author can republish the web page hosting the document and interact with the
proposed revision. As shown in \secref{}, this is an important validation step that can reveal errors in the
generated expression. If the interactions look reasonable, the author can approve the new document state; this
is the primary successful path through the workflow and returns the system to the entry state where it is
waiting for another top-level interaction from the author. Otherwise, the author rejects the proposed change
and returns to the entry state without any change to the document.

>>>>

\item \textbf{High-level Authoring workflow.} The system then enters the human-in-the loop authoring workflow
shown in Figure~\ref{fig:interpretation-agent}, where the author interacts with the \InterpretationAgent. The
system waits for the author to select a fragment of text $s$ to interpret (perhaps previously highlighted by
the \SuggestionAgent). The system then attempts to generate a candidate Fluid expression $e$ using the
closed-loop synthesis step (3) below. If code synthesis succeeds with an expression $e$, the system proceeds
to the manual validation step (4) below. If the synthesis step fails with no expression, no remedial action is
possible; this is considered an unsuccessful path through the workflow and returns the system to the entry
state. Otherwise the synthesis step produces an expression $e$ which evaluates to a mismatched string $s' \neq
s$ outcome, and the user can choose to manually abort and return to the entry state, or optionally to
\emph{revise the goal}, replacing $s$ with $s'$ in the target document and retaining $e$ as the candidate
expression. This is intended to cover the situation where the author has made a claim which is
\emph{incorrect}, and the dataset and surrounding natural language have led the LLM to synthesise an
expression which generates a different value from the one specified by the user.

@@

\item \textbf{Manual validation step.} Once a candidate expression has been generated, the system replaces the
selected substring $s$ with the interpolation expression \kw{\{$e$\}}, creating a new (but only tentative)
document configuration. The author can republish the web page hosting the document and interact with the
proposed revision. As shown in \secref{evaluation}, this is an important validation step that can reveal
errors in the generated expression. If the interactions look reasonable, the author can approve the new
document state; this is the primary successful path through the workflow and returns the system to the entry
state where it is waiting for another top-level interaction from the author. Otherwise, the author rejects the
proposed change and returns to the entry state without any change to the document.

### 2025-11-21 paper/sec/introduction.tex
Fix borked refs.
Break down by RQ1 and RQ2.

\paragraph{Contributions.} Our specific contributions are as follows. (\secref{conclusion:future-work}). We
leave implementing a full Copilot-like authoring plugin for an IDE such as VSCode or Cursor for future work.

\begin{itemize}
\item A proof-of-concept LLM-based tool for iteratively transforming a preexisting opaque document and
associated data set into a transparent, data-driven counterpart (\secref{authoring-workflow});
\item A summary of the natural language idioms we have studied (\secref{nl-idioms}) and an empirical
evaluation of how well state-of-the-art models are able to solve the associated interpretation and code
synthesis problems (\secref{evaluation}).
\end{itemize}

>>>>

\paragraph{Contributions.} Our specific contributions are as follows. We leave implementing a full
Copilot-like authoring plugin for an IDE such as VSCode or Cursor for future work (\secref{conclusion}).

\begin{itemize}
\item A proof-of-concept LLM-based tool for iteratively transforming a preexisting opaque document and
associated dataset into a transparent, data-driven counterpart (\secref{authoring-workflow});
\item A summary of the natural language idioms we have studied (\secref{nl-idioms}) and an empirical
evaluation of how well state-of-the-art models are able to solve the associated interpretation and code
synthesis problems (\secref{evaluation}).
\end{itemize}
